FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-09 ess_marmo.tex
Update on Overleaf.

This paper proposes to use Flexible Image Transport
System\footnote{https://fits.gsfc.nasa.gov/}
(FITS) format \citep{fitsorig,fitsver3} in planetary surface investigations,
and describes how FITS metadata can easily be inserted in the
PDS4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml}
metadata distribution model.
FITS is one of the standard implemented formats in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html}
(VO), therefore making obvious its connection to the Planetary Virtual
Observatory initiative and the planetary Table Access Protocol (EPN-TAP)
\citep{ERARDepntap}.
FITS format is open and flexible, largely implemented in open and efficient
processing tools.
Those reasons make FITS format appropriate for large amounts of raster data
processing.
The goal of this approach is to make easier data mining and reprocessing
in planetary surface investigations, promoting general software based on those
standards.
The option to use FITS within the planetary domain could potentially
homogenize methods from acquisition, to visualization, while
giving chances to optimize data processing.

@@

\subsubsection{OMEGA Spectra and Geometry Data}
As stressed above, having planetary data distributed in FITS at any processing level is
a way to avoid multiple conversions in the processing chain and simplify reprocessing.
In planetary science a first spatial calibration data often consists in providing coordinates
on specific points on the detector (the center of the pixel, one or more pixel corners, etc.).
To support such look-up table coordinate representation the \texttt{TAB} algorithm
has been defined in \citet{wcsspectral}.
In the \texttt{TAB} algorithm, coordinates are listed in a coordinate array, an indexing vector
can be used to address coordinate array elements.
Coordinates can then be sampled more or less coarsely depending on the behavior of the spatial
reconstruction.
Also, when the field is only partially covered by the planetary surface, coordinate array
dimensions can be significantly reduced. 
TAB projection is implemented in the Calabretta WCSLIB\footnote{http://www.atnf.csiro.au/people/mcalabre/WCS/wcslib/},
available in all major linux distribution.

@@

OMEGA rasters are equally sampled all over the field: we use them to exemplify
TAB description without indexing, then we only use spatial information on the center
of the pixel. 
The output result can be downloaded at the address ...

@@

\subsubsection{Akatsuki Imaging Data}

\subsubsection{HiRISE Mosaics}

@@

\section{Discussion and Perspectives}
\label{sec:disc}
FITS discussion in astronomy\citep{THOMAS2015133} 
let's astronomy learn from planetary sciences too: accent must be put on interoperability
.... blabla
Planetary surface is lacking open and modular software for massive processing and can benefit
of astronomy experience. In addition to that WCS projection scheme is simpler than the
historical one used in planetary applications. 
Archiving flexibility and data model are today approached by PDS4 and EPN-TAP efforts.
Reducing custom data format conversions during the processing chain but
allowing easy conversion for final products following the needed application.

>>>>

This paper proposes to use Flexible Image Transport
System\footnote{https://fits.gsfc.nasa.gov/}
(FITS) format \citep{fitsorig,fitsver3} in planetary surface investigations,
and describes how FITS metadata can easily be inserted in the
PDS4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml}
metadata distribution model.
FITS is one of the standard implemented formats in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html}
(VO), therefore making obvious its connection to the Planetary Virtual
Observatory initiative and the planetary Table Access Protocol (EPN-TAP)
\citep{ERARDepntap}.
FITS format is open and flexible, largely implemented in open and efficient
processing tools.
Those reasons make FITS format appropriate for large amounts of raster data
automatic processing.
The goal of this approach is to make easier data mining and reprocessing
in planetary surface investigations, promoting general software based on those
standards.
The option to use FITS within the planetary domain could potentially
homogenize methods from acquisition, to visualization, while
giving chances to optimize data processing.

@@

\subsubsection{OMEGA Spectra and Geometry Data}
As stressed above, having planetary data distributed in FITS at any processing level is
a way to avoid multiple conversions in the processing chain and simplify reprocessing.
In planetary science a first spatial calibrated product often consists in providing coordinates
on specific points on the detector (the center of the pixel, one or more pixel corners, etc.).
To support such look-up table coordinate representation the \texttt{TAB} algorithm
has been defined in \citet{wcsspectral}.
In the \texttt{TAB} algorithm, coordinates are listed in a coordinate array, an indexing vector
can be used to address coordinate array elements.
Coordinates can then be sampled more or less coarsely depending on the behavior of the spatial
reconstruction.
Also, when the field is only partially covered by the planetary surface, coordinate array
dimensions can be significantly reduced. 
TAB projection is implemented in the Calabretta WCSLIB\footnote{http://www.atnf.csiro.au/people/mcalabre/WCS/wcslib/},
available in all major linux distribution.

@@

OMEGA rasters are equally sampled all over the field: we use them to exemplify
TAB description without indexing, then we only use spatial information on the center
of the pixel. 
The output result can be downloaded at the address https://github.com/cmarmo/convertofits/releases/download/v1.0-beta/orb0413\_1.IR1.fits.gz together with the python script
used for conversion\footnote{https://github.com/cmarmo/convertofits/blob/master/omega2fits.py}.

@@

\subsubsection{Akatsuki Imaging Data}
The Venus Climate Orbiter Akatsuki\footnote{http://akatsuki.isas.jaxa.jp/en/} is a JAXA
spacecraft dedicated to the study of the Venus atmosphere.
Akatsuki imaging data are already distributed in FITS format.
Geometry information is distributed in two different cubes of data, containing respectively 
values at the center of the pixel and at the lower left corner of the pixel.
Tabular coordinate representation allows us to gather geometry in one table together with
the described raster.
In addition to that, in Akatsuki images Venus is often far from completely filling the field:
tabular representation avoids to store unnecessary data. 
In that case, coordinate indexes are necessary to identify which pixels contains the spatial
information.
The output result can be downloaded at the address https://github.com/cmarmo/convertofits/releases/download/v1.0-beta/ir1\_20160415\_070351\_097\_l2b\_v10\_out.fit.gz
together with the python script
used for conversion\footnote{https://github.com/cmarmo/convertofits/blob/master/akatsukil2b.py}.

Again, we used \texttt{wcsware} to validate the WCS structure and to test pixel to world /
world to pixel coordinate conversion.
Even if TAB projection is not fully implemented in FITS visualization tools, this representation
has the advantage to store the detector geometrical information together with physical
quantities measured by the detector itself in a standardized way.

@@

\section{Discussion and Perspectives}
\label{sec:disc}
FITS discussion in astronomy\citep{THOMAS2015133} 
let's astronomy learn from planetary sciences too: accent must be put on interoperability
.... blabla
Planetary surface is lacking open and general software for massive processing and can benefit
of astronomy experience. In addition to that WCS projection scheme is simpler than the
historical one used in planetary applications. 
Archiving flexibility and data model are today approached by PDS4 and EPN-TAP efforts.
Reducing custom data format conversions during the processing chain but
allowing easy conversion for final products following the needed application.
